### 2022-01-25 paper/manuscript.Rmd
Standardization Attempt 2
Second attempt at writing the piece on standardization.

Penalized regression analyses often require the scales of the predictors to be equalized ([@tibshirani1996; @lee2015note]). This is because $\lambda$, the regularization parameter, penalizes the magnitude of the coefficients that are dependent on the scales of the predictors. This can lead to the unfair penalization of larger coefficients that do not necessarily reflect importance ([@zou2006adaptive]). To clarify, a regression parameter $\beta$ can be interpreted as the amount outcome $\mathbf{Y}$ increases as predictor $\mathbf{X}$ increases with 1 unit. This means that as $\mathbf{X}$ is reduced by a factor of 10, its regression coefficient is increased by a factor of 10. So suppose the desire to predict running speed by a person`s height. If height is measured in kilometers, than a one unit increase in height, which is a one kilometer increase, will likely lead to a drastic increase in running speed. In penalized regression techniques, this high value for $\beta$ would then dominate the penalty function, leaving coefficients of other predictors unregularized.

Thus, to equalize predictor scales, a widely used technique is standardization ([@gelman2008scaling]). Standardization sets the mean of all predictors to 0 and their standard deviation to 1. It does so by Z-transforming all predictors as follows:

$$
\mathbf{Z} = \frac{\mathbf{X}- \bar{\mathbf{X}}}{\mathbf{s_X}}
$$
where $\bar{\mathbf{X}}$ is a vector of predictor means and $\mathbf{s_X}$ is a vector of standard deviations of the predictors. `brma()` standardizes the predictors by default on which the model is estimated. The function returns the estimated parameters back in their original scales. This is achieved for the intercept by applying:

$$
b_0 = b_{0Z} - \mathbf{b}_Z\frac{\bar{\mathbf{X}}}{\mathbf{s}_X} \\
$$
where $b_0$ is the intercept, $b_{0Z}$ is the intercept for the standardized predictors and $\mathbf{b_Z}$ is the vector of regression coefficients for the standardized predictors. The regression coefficients are returned in their original scale by applying:

@@

There is some debate regarding the standardization of binary/dummy predictors in penalized regression ([@lee2015note]). Some suggest to always standardize dummy variables ([@tibshirani1997lasso]). However, standardizing binary predictors increases the absolute maximum values the predictor can take on which in some cases can make its coefficient more sensitive to the penalty function ([@harrell2001regression]).

In any case, the default standardization option in `brma()` can be overwritten entirely by setting `standardize = False`. `brma()` also allows the user to standardize only some of the predictors. This is done by constructing a vector of means and a vector of standard deviations. These vectors contain the means and standard deviations of the predictors in their original scales, except for the predictors that **should not** be standardized. For the latter, a mean of 0 and a standard deviation of 1 must be passed. The predictors that need to be standardized should then be standardized manually. Finally, the mean and standard deviation vectors are passed to the `standardize` argument as follows: `standardize = list(center = MeanVector, scale = SdVector)`.

>>>>

Penalized regression analyses often require the scales of the predictors to be equalized ([@tibshirani1996]). This is because $\lambda$, the regularization parameter, penalizes the magnitude of the coefficients that are dependent on the scales of the predictors. This can lead to the unfair penalization of larger coefficients that do not necessarily reflect importance ([@lee2015note]). To clarify, a regression parameter $\beta$ can be interpreted as the amount outcome $\mathbf{Y}$ increases as predictor $\mathbf{X}$ increases with 1 unit. This means that as $\mathbf{X}$ is reduced by a factor of 10, its regression coefficient is increased by a factor of 10. So suppose the desire to predict running speed by a person`s height. If height is measured in kilometers, than a one unit increase in height, which is a one kilometer increase, will likely lead to a drastic increase in running speed. In penalized regression techniques, this high value for $\beta$ would then dominate the penalty function, leaving coefficients of other predictors unregularized.

To equalize predictor scales, a widely used technique is standardization ([@gelman2008scaling]). Standardization sets the mean of all predictors to 0 and their standard deviation to 1. It does so by Z-transforming all predictors as follows:

$$
\mathbf{Z} = \frac{\mathbf{X}- \bar{\mathbf{X}}}{\mathbf{s_X}}
$$
where $\bar{\mathbf{X}}$ is a vector of predictor means and $\mathbf{s_X}$ is a vector of standard deviations of the predictors. `brma()` standardizes the predictors by default on which the model is estimated. The function returns the estimated parameters back to their original scales. This is achieved for the intercept by applying:

$$
b_0 = b_{0Z} - \mathbf{b}_Z\frac{\bar{\mathbf{X}}}{\mathbf{s}_X} \\
$$
where $b_0$ is the intercept, $b_{0Z}$ is the intercept for the standardized predictors and $\mathbf{b_Z}$ is the vector of regression coefficients for the standardized predictors. The regression coefficients are returned to their original scale by applying:

@@

It is not always necessary or desirable to standardize predictors, however. The most straightforward instance is when predictors are already standardized or on the same scale for different reasons. In these cases standardization does not make scales more equal, nor the penalization more fair.

Also, if the desire is to directly compare current research outcomes with prior research outcomes that did not standardize predictors, standardization in the current research would make penalization unequal across studies.

Finally, standardization should be thought over when dealing with binary/dummy predictors ([@alkharusi2012categorical]). Some suggest to always standardize binary predictors ([@tibshirani1997lasso]). This makes that, irrespective of initial scaling, the binary predictor will be on the same scale as the continuous standardized predictors ([@gelman2008scaling]). However, standardizing binary predictors may decrease model interpretability ([@wissmann2007role]). Consider a regression problem including only a binary predictor $X$ that takes on values 0 and 1 predicting outcome $Y$. The intercept represents $\mathbb{E}(Y | X = 0)$ and the regression coefficient represents $\mathbb{E}(Y | X = 1)$, or the difference in means on $Y$ between conditions ([@alkharusi2012categorical]). By standardizing binary predictors, they lose 0 as a value to take on and both the intercept and its regression coefficient have no clear interpretation anymore, especially in multivariate cases ([@wissmann2007role]).

Also note that there are multiple coding schemes for categorical predictors in penalized regression, but that the coding choice can affect model accuracy and interpretation ([@chiquet2016coding; @detmer2020note]). The standard option is to dummy code categorical predictors and standardize the dummies afterwards, dependent on considerations mentioned above.

For all cases where standardization is undesired, the default standardization option in `brma()` can be overwritten entirely by setting `standardize = False`. `brma()` also allows the user to standardize only some of the predictors. This is done by constructing a vector of means and a vector of standard deviations of the predictors in their original scales, except for the predictors that **should not** be standardized. For the latter, a mean of 0 and a standard deviation of 1 must be passed. The predictors that need to be standardized should then be standardized manually. Finally, the mean and standard deviation vectors are passed to the `standardize` argument as follows: `standardize = list(center = MeanVector, scale = SdVector)`.
